---
title: "Assignment_2.1_dataRetrieval"
author: "Julia Parish"
date: "10/13/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dataRetrieval)
library(tidyverse)
library(here)
library(lubridate)

```

# Assignment 2.1 dataRetrieval
1. Add the Santa Paula Creek gauge to exercise we did in class
2. What is the best way to query multiple sites using this R package?
3. Create a plot showing any stream flow response to the rain of last night (10/04). Determine at what time did the stream flow peaked (if any) at the two locations?

## 2.1.1 Add the Santa Paula Creek gauge to exercise we did in class
```{r combined waterways}

# Ventura River and Santa Paula Creek discharge rates
sitenumber <- c("11118500", "11113500")

parameterd <- "00060" # discharge

startdate <- "2019-10-01"

enddate <- "2020-10-05"

#Raw discharge data:
combined_discharge <- readNWISdv(sitenumber,
                        parameterd,
                        startdate,
                        enddate) %>% 
                      ## rename column x_00060_00003 to discharge
                      rename(discharge = X_00060_00003, code = X_00060_00003_cd) %>% 
                      mutate(site_name = if_else(site_no == "11118500", "Ventura River",
                                   if_else(site_no == "11113500", "Santa Paula Creek", ""))) %>%  
                      relocate(site_name, .before = "site_no")

```

```{r discharge plot}

discharge_plot <- combined_discharge %>% 
                  ggplot(aes(color = site_name)) +
                  geom_line(aes(x = Date, y = discharge)) +
                  labs(title = "Discharge Rates of Ventura River & Santa Paula Creek, California",
                       x = "Date and Time (PST)",
                       y = "Discharge Rate in cubic feet",
                       color = "Site") +
                  theme_minimal()

discharge_plot


```


## 2.1.2 What is the best way to query multiple sites using this R package?

* Assigning sites as vectors allows you to query the same variables and dates at the same time. 

```{r mean max}
statCd <- c("00001","00003")  # Mean and maximum

maxmeanflow <- readNWISdv(sitenumber,
                   parameterd,
                   startdate, 
                   enddate,
                   statCd = statCd)
  
maxmeanflow <- renameNWISColumns(maxmeanflow) %>% 
                # add a new column that list the site name              
                mutate(site_name = if_else(site_no == "11118500", "Ventura River",
                                   if_else(site_no == "11113500", "Santa Paula Creek", ""))) %>%  
                relocate(site_name, .before = "site_no") # move new site name column to before site_no# rename column names using the renameNWISColumns function

#names(attributes(maxmeanflow))


```

```{r maxmeanflow plot}

maxmeanflow_plot <- maxmeanflow %>% 
                  ggplot(aes(color = site_name)) +
                  geom_line(aes(x = Date, y = Flow)) +
                  labs(title = "Flow Rates of Ventura River & Santa Paula Creek, California",
                       subtitle = "Max and Mean Rates",
                       x = "Date and Time (PST)",
                       y = "Flow Rate in cubic feet",
                       color = "Site") +
                  theme_minimal()

maxmeanflow_plot
```



## 2.1.3 Create a plot showing any stream flow response to the rain of last night (10/04). Determine at what time did the stream flow peaked (if any) at the two locations?

```{r read in site info}

startdate = as.Date("2021-10-04T00:00")
enddate = as.Date("2021-10-05T24:00")

spv_info <- readNWISsite(sitenumber)
  
combined_flow = readNWISuv(
                  siteNumber = c("11118500", "11113500"),
                  parameterCd = "00060",
                  startDate = startdate,
                  endDate = enddate,
                  tz = "PST")

combined_flow <- renameNWISColumns(combined_flow) %>% 
                rename(flow = Flow_Inst) %>% # rename column names using the renameNWISColumns function
                # add a new column that list the site name              
                mutate(site_name = if_else(site_no == "11118500", "Ventura River",
                                   if_else(site_no == "11113500", "Santa Paula Creek", ""))) %>%  
                relocate(site_name, .before = "site_no") # move new site name column to before site_no
  

```

```{r filter sites}
# filter by site

# ventura river flow data
ventura_flow <- filter(combined_flow, site_no == "11118500")

# santa paula creek flow data
spc_flow <- filter(combined_flow, site_no == "11113500")


```


```{r combined plot}

stormevent_plot <- combined_flow %>% 
                  ggplot(aes(color = site_name)) +
                  geom_line(aes(x = dateTime, y = flow)) +
                  # geom_vline(xintercept = max(spc_flow$flow), 
                  #            color = "#444343", size = 1.1) +
                  annotate("text", 
                       label = "Max = 0.70\n10/05 3:30AM", 
                       x = as.POSIXct("2021-10-05 04:00:00"), 
                       y = 0.75, 
                       color = "black",
                       size = 3) +
                  # geom_vline(xintercept = max(ventura_flow$flow), 
                  #            color = "#444343", size = 1.1) +
                  annotate("text", 
                       label = "Max = 1.08\n10/05 12:25AM", 
                       x = as.POSIXct("2021-10-05 01:30:00"), 
                       y = 0.98, 
                       color = "black",
                       size = 3) +
                  labs(title = "Flow Rates of Ventura River & Santa Paula Creek, California",
                       subtitle = "Storm Event Occurred on October 04, 2021",
                       x = "Date and Time (PST)",
                       y = "Flow Rate in cubic feet",
                       color = "Site") +
                  theme_minimal()

stormevent_plot
```



